---
title: "Lab 12 - Smoking during pregnancy"
author: "Fiona Wang"
date: "Apr 2 2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r load-data}
data <- ncbirths
```

```{r checktype}
glimpse(data)
```

```{r outlier}
boxplot.stats(data$fage)$out
boxplot.stats(data$mage)$out
boxplot.stats(data$weeks)$out
boxplot.stats(data$visits)$out
boxplot.stats(data$gained)$out
boxplot.stats(data$weight)$out
```

Looking at this tibble, there are a couple numerical variables: father's age, mother's age, length of pregnancy in weeks, number of hospital visits during pregnancy, weight gained by mother during pregnancy in pounds, and weight of the baby at birth in pounds. The others are all categorical. For the numerical variables, all of them have outliers. 

### Exercise 1
There are 1000 cases in this dataset. THe cases are the information on birth record in the state of North Carolina from 2004. 

### Exercise 2
```{r filter}
ncbirths_white <- data %>% 
  filter(whitemom == "white")
ncbirths_white %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

The mean weights of the white babies is 7.25 pounds. 

### Exercise 3
I think it meets the criteria necessary to conduct simulation-based inference. It is a large enough sample size, so we could simulate samples easily. The observations are independent, I think. 

### Exercise 4
```{r generatebootstrap}
set.seed(1232)
boot_df <- ncbirths_white %>% 
  specify(respons = weight) %>% 
  hypothesize(null = "point", mu = 7.43) %>% 
  generate(reps = 10000, type  = "bootstrap") %>% 
  calculate(stat = "mean")
glimpse(boot_df)
```

Now, let's plot the graph.
```{r plotdistr}
ggplot(data = boot_df, mapping = aes(x = stat)) +
  geom_histogram(binwidth = 0.001) +
  labs(title = "Null Distribution") +
  geom_vline(xintercept = 7.43, color = "red")
```

It seems to be pretty normally distributed (or a little left skewed), and centered around 7.43. 
```{r pvalue}
boot_df %>% 
  summarize(lower = quantile(stat, .025),
            uppter = quantile(stat, .975))
boot_df %>% 
  summarize(p_value = mean(abs(stat - 7.43) >= abs(7.25 - 7.43)))
```

It seems like the p-value is 0.0007. There is very few instances in the null distribution that we see a difference as large as the one we observed in the current dataset. The null's mean is 7.43, the observed mean is 7.25, which is a 0.18 difference. In our 10,000 bootstrap results, there were few instances where they had a bigger than 0.18 difference. So, it is very unlikely to see an instance as extreme as ours. Thus, we conclude that the weight of White babies in NC in 2004 is significantly lighter than the White babies in 1995. 

### Exercise 5
```{r boxplot}
boxplot(data$weight ~ data$habit, data = data,
        col = c("skyblue", "lightgreen"))
```

From the boxplot, we see that the weight of the babies whose mothers were nonsmokers on average was heavier than that of the babies whose mothers were smokers. 

### Exercise 6
```{r filtermissing}
ncbirths_clean <- data %>% 
  filter(!is.na(habit), !is.na(weight))
```

It seems like there is only one missing value.

### Exercise 7
```{r compareweight}
ncbirths_clean %>% 
  group_by(habit) %>% 
  summarize(mean_weight = mean(weight))
```

### Exercise 8
Write the hypotheses testing to see if the difference is significant.
H0: the average weight of babies born to smoking mothers and the average weight of babies born to non-smoking mothers are not significantly different. 
HA: the average weight of babies born to smoking mothers are significantly lighter than babies born to non-smoking mothers. 

### Exercise 9
```{r t-test}
t.test(weight ~ habit, data = ncbirths_clean)
```

According to the t-test, the p-value is 0.019 < 0.05. This means that it is very unlikely to observe a data as extreme as ours, so the two groups significantly differ from each other in the average weight of the babies. The average weight of babies of non-smoker mothers is significantly heavier than the average weight of babies of smoker mothers. 

### Exercise 10
The t-test already gives us the 95% CI around the difference: [.052, .580]. 
```{r CIdiff}
set.seed(123)
boot_df2 <- ncbirths_clean %>%
  specify(weight ~ habit) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("nonsmoker", "smoker"))
ci <- boot_df2 %>%
  get_confidence_interval(level = 0.95, type = "percentile")
print(ci)
```

From the bootstrap the confidence interval is a little different from what the t-test gave me. [.059, .580].

### Exercise 11
From my experience, I would use median cut-offs because that's what I did when determining low and high SES groups. We can divide mothers equally into two groups. 
```{r cutoff}
median_age <- median(data$mage, na.rm = TRUE)
median_age
```

The median age of mothers is 27, and that's the cutoff if I were to determine it. For mothers who are younger than 27, they are younger mothers, for mothers who are 27 or older, they are mature mothers. 

### Exercise 12
Hypothesis:
H0: There is no true difference in the proportion of low birth weight babies between younger and mature mothers.
HA: The proportion of low birth weight babies is higher for mature mothers. 

```{r boot3}
#str(data$mature)
#str(data$lowbirthweight)
ncbirths_clean <- ncbirths_clean %>%
  mutate(lowbirthweight_num = if_else(lowbirthweight == "low", 1, 0))
obs_diff <- ncbirths_clean %>%
  group_by(mature) %>%
  summarize(prop_low = mean(lowbirthweight_num), .groups = "drop") %>%
  pivot_wider(names_from = mature, values_from = prop_low) %>%
  mutate(diff = `mature mom` - `younger mom`) %>%
  pull(diff)
set.seed(123)
boot_df3 <- ncbirths_clean %>%
  specify(lowbirthweight_num ~ mature) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("younger mom", "mature mom"))
p_value3 <- boot_df3 %>%
  summarize(p_value = mean(stat >= obs_diff)) %>%
  pull(p_value)
print(p_value3)
```
The observed difference is 2% in the low weight proportion. However, the p-value is 0.07, which indicate that the result is not significant. So this 2% is likely due to chance. Fail to reject the null. There is no significant difference in the proportion of low birth weight babies in mature or not mature mom. 

### Exercise 13
```{r CI3}
ci <- boot_df3 %>%
  get_confidence_interval(level = 0.95, type = "percentile")
print(ci)
```
Based on the sample of mothers in this study, we are 95% confident that the true difference in the proportion of low birthweight babies between mature mothers and younger mothers is between 8.4% lower and 3.7% higher for mature mothers. Since this CI includes 0, it means that there is not enough evidence to conclude that low birth weight percentage is higher or lower for mature mothers. 