---
title: "Lab 12 - Smoking during pregnancy"
author: "Fiona Wang"
date: "Apr 2 2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r load-data}
data <- ncbirths
```

```{r checktype}
glimpse(data)
```

```{r outlier}
boxplot.stats(data$fage)$out
boxplot.stats(data$mage)$out
boxplot.stats(data$weeks)$out
boxplot.stats(data$visits)$out
boxplot.stats(data$gained)$out
boxplot.stats(data$weight)$out
```

Looking at this tibble, there are a couple numerical variables: father's age, mother's age, length of pregnancy in weeks, number of hospital visits during pregnancy, weight gained by mother during pregnancy in pounds, and weight of the baby at birth in pounds. The others are all categorical. For the numerical variables, all of them have outliers. 

### Exercise 1
There are 1000 cases in this dataset. THe cases are the information on birth record in the state of North Carolina from 2004. 

### Exercise 2
```{r filter}
ncbirths_white <- data %>% 
  filter(whitemom == "white")
ncbirths_white %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

The mean weights of the white babies is 7.25 pounds. 